Outbreak Threshold Detection

### 2023-03-06 notebooks/Julia/simulation-detection.qmd
Convert back to Gillespie Direct Using BenchmarkTools indicated that Gillespie Direct was quicker than my implementation of Tau-Leaping or SDEProblem, and solves the main issue I was having of jumps going below 0. Look into performance again if necessary

```{julia}
using JumpProcesses, StochasticDiffEq, DifferentialEquations, LinearAlgebra
using Catalyst, Random, Statistics, DataFrames, CairoMakie, ProgressMeter

@@

```{julia}
function rate(out, u, p, t)
    out[1] = p[1] * u[1] * u[2] / (u[1] + u[2] + u[3])           # β * S * I
    out[2] = p[2] * u[2]                      # ν * I
    # out[3] = p[3] * u[1]                    # μ * N
    # out[4] = p[3] * u[1]                    # μ * S
    # out[5] = p[3] * u[2]                    # μ * I
    # out[6] = p[3] * u[3]                    # μ * R
    nothing
end
```

```{julia}
c = zeros(3, 2)
# S + I --> I
c[1, 1] = -1        # S -> S - 1
c[2, 1] = 1         # I -> I + 1

# I --> R
c[2, 2] = -1        # I -> I - 1
c[3, 2] = 1         # R -> R + 1

# S --> 2X
# c[1, 3] = 1         # S -> S + 1

# # S --> 0X
# c[1, 4] = -1        # S -> S - 1

# # I --> 0Y
# c[2, 5] = -1        # I -> I - 1

# # R --> 0Z
# c[3, 6] = -1        # R -> R - 1

function change!(du, u, p, t, counts, mark)
    mul!(du, c, counts)
    nothing
end
```

```{julia}
rj = RegularJump(rate, change!, 2)
```

```{julia}
tmax = 100.0
δt = 0.001
tspan = (0.0, tmax)
tlength = length(tspan[1]:δt:tspan[2])
Random.seed!(1234);
```

```{julia}
# R₀ = 2
# γ = 1/8
μ = 0.001
# β = R₀ * (γ + μ)
# ζ = 0.001
# p = (β, γ, μ, ζ)

u₀ = [1000.0, 50.0, 0.0]

β = 0.1/1000.0
γ = 0.01
μ = 0.001
p = (β, γ, μ)

```

```{julia}
prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Direct(), rj)
sol = solve(jump_prob, TauLeaping(); dt = δt)
sol_array = zeros(4, length(sol))
sol_array[1:3, :] = Array(sol)
sol_array[4, :] = sum(sol_array[1:3, :], dims = 1)
```

```{julia}
f = Figure()
ax = Axis(
    f[1, 1],
    xlabel = "Time",
    ylabel = "Number"
)

lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "S(t)")
lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "I(t)")
lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "R(t)")
lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "N(t)")

@@

f
```

>>>>

```{julia}
using JumpProcesses, StochasticDiffEq, DifferentialEquations, LinearAlgebra
using Catalyst, Random, Statistics, DataFrames, CairoMakie, ProgressMeter
using BenchmarkTools

@@

```{julia}
tmax = 2.0
δt = 0.1/365
tspan = (0.0, tmax)
tlength = length(tspan[1]:δt:tspan[2])
Random.seed!(1234);

u₀ = [999.0, 10.0, 0.0, 0.0]

μ = 1
β = 1000
σ = 365/8
γ = 365/5
p = (β, σ, γ, μ)
```

```{julia}
expose_rate(u, p, t) = p[1] * u[1] * u[2] / (u[1] + u[2] + u[3] + u[4])  # β * S * I / N
function expose_affect!(integrator)
    integrator.u[1] -= 1         # S -> S - 1
    integrator.u[2] += 1         # E -> E + 1
    nothing
end
expose_jump = ConstantRateJump(expose_rate, expose_affect!)

infec_rate(u, p, t) = p[2] * u[2]   # σ * E
function infec_affect!(integrator)
    integrator.u[2] -= 1         # E -> E - 1
    integrator.u[3] += 1         # I -> I + 1
    nothing
end
infec_jump = ConstantRateJump(infec_rate, infec_affect!)

recov_rate(u, p, t) = p[3] * u[3]   # γ * I
function recov_affect!(integrator)
    integrator.u[3] -= 1         # I -> I - 1
    integrator.u[4] += 1         # R -> R + 1
    nothing
end
recov_jump = ConstantRateJump(recov_rate, recov_affect!)

birth_rate(u, p, t) = p[4] * (u[1] + u[2] + u[3] + u[4]) # μ * N
function birth_affect!(integrator)
    integrator.u[1] += 1         # S -> S + 1
    nothing
end
birth_jump = ConstantRateJump(birth_rate, birth_affect!)

S_death_rate(u, p, t) = p[4] * u[1] # μ * S
function S_death_affect!(integrator)
    integrator.u[1] -= 1        # S -> S - 1
end
S_death_jump = ConstantRateJump(S_death_rate, S_death_affect!)

E_death_rate(u, p, t) = p[4] * u[2] # μ * E
function E_death_affect!(integrator)
    integrator.u[2] -= 1        # E -> E -1
end
E_death_jump = ConstantRateJump(E_death_rate, E_death_affect!)

I_death_rate(u, p, t) = p[4] * u[3] # μ * I
function I_death_affect!(integrator)
    integrator.u[3] -= 1        # I -> I - 1
end
I_death_jump = ConstantRateJump(I_death_rate, I_death_affect!)

R_death_rate(u, p, t) = p[4] * u[4] # μ * R
function R_death_affect!(integrator)
    integrator.u[4] -= 1        # R -> R - 1 
end
R_death_jump = ConstantRateJump(R_death_rate, R_death_affect!)

gillespie_jumps = [
    expose_jump, infec_jump, recov_jump, birth_jump, S_death_jump, E_death_jump,
    I_death_jump, R_death_jump
]
```

```{julia}
prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Direct(), gillespie_jumps...)
sol = solve(jump_prob, SSAStepper())
sol_array = zeros(5, length(sol))
sol_array[1:4, :] = Array(sol)
sol_array[5, :] = sum(sol_array[1:4, :], dims = 1)
```

```{julia}
fig = Figure()
ax = Axis(
    fig[1, 1],
    xlabel = "Time",
    ylabel = "Number"
)

lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "S(t)")
lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "E(t)")
lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "I(t)")
lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "R(t)")
lines!(ax, sol.t, sol_array[5, :], linewidth = 3, label = "N(t)")

@@

fig
```
